---
title: "Course reproducible-research Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.Code for reading in the dataset and/or processing the data

```{r reading}
library(knitr)
library(readr)
activity <- read_csv("repdata_data_activity/activity.csv",
col_types = cols(date = col_date(format = "%Y-%m-%d"),
steps = col_number()))
dim(activity)
summary(activity)
```

## 2.Histogram of the total number of steps taken each day

```{r histogram, echo=TRUE}
library(dplyr)
steps_day <- activity %>% group_by(date)%>%summarise(tot_steps=sum(steps,na.rm=TRUE))
library(ggplot2)
qplot(tot_steps,data=steps_day, main = "Total steps by day",  xlab = "day", ylab = "Steps")
```


## 3.Mean and median number of steps taken each day
```{r mean, echo=TRUE}
library(dplyr)
m_day <- activity %>% group_by(date)%>%summarise(mean_steps=mean(steps,na.rm=TRUE), median_steps=median(steps,na.rm=TRUE))
m_day
```

## 4.Time series plot of the average number of steps taken
```{r time series, echo=TRUE}
library(ggplot2)
qplot(date,mean_steps,data=m_day,geom=c("point","line"), main = "Average number of steps taken",  xlab = "day", ylab = "mean steps")
```

## 5.The 5-minute interval that, on average, contains the maximum number of steps
```{r interval, echo=TRUE}
library(dplyr)
m_interval<- activity %>% group_by(interval) %>% summarise(average=mean(steps,na.rm=TRUE))
max_numb_steps<- filter(m_interval,average==max(average))
max_numb_steps
```

## 6.Code to describe and show a strategy for imputing missing data
### There are 8 dates with no data points. I'm going to impute missing value with average of steps by weekdays and interval 
```{r missing, echo=TRUE}
library(lubridate)
library(dplyr)
activity<-activity %>% mutate(wday_= wday(date))
imp<-activity %>% group_by(wday_, interval) %>% summarise(ave_wday=mean(steps,na.rm=TRUE))
act_imp<-left_join(activity, imp, by =c("wday_", "interval"))
act_imp<-act_imp %>% mutate(steps_=ifelse(is.na(steps), ave_wday,steps))
summary(act_imp)
```
## 7.Histogram of the total number of steps taken each day after missing values are 
```{r impute, echo=TRUE}
library(dplyr)
steps_day_imp <- act_imp %>% group_by(date)%>%summarise(tot=sum(steps_))
library(ggplot2)
qplot(tot,data=steps_day_imp, main = "Total steps by day",  xlab = "day", ylab = "Steps")
```

## 8.Panel plot comparing the average number of steps taken per 5-min interval across weekends
### I'm going to impute missing value with average of steps by weekdays and interval 
```{r wdays, echo=TRUE}
library(lubridate)
library(dplyr)
library(ggplot2)
activity<-activity %>% mutate(wday_= wday(date),wdya_l=ifelse(wday_==7 | wday_==1, "weekend","weekday"))
imp<-activity %>% group_by(wday_, interval) %>% summarise(ave_wday=mean(steps,na.rm=TRUE))
act_imp<-left_join(activity, imp, by =c("wday_", "interval"))
act_imp<-act_imp %>% mutate(steps_=ifelse(is.na(steps), ave_wday,steps))
m_day_wd <- act_imp %>% group_by(interval,wdya_l)%>%summarise(mean_steps_wday=mean(steps_,na.rm=TRUE))
qplot(interval,mean_steps_wday,data=m_day_wd,facets=~wdya_l,color=wdya_l,geom=c("line"), main = "average of steps per 5-minute interval across weekdays and weekends",  xlab = "interval", ylab = "mean steps")
```



